HUME: Human UCCA-Based Evaluation of Machine Translation

### 2016-09-13 paper/main.tex
My changes for camera ready

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence. We 
are thus avoiding the semantic annotation of machine-generated text,
which is often garbled or semantically unclear.
This also allows the re-use of the source semantic annotation for
measuring the quality of different translations of the same source sentence,
and avoids relying on possibly suboptimal reference translations.

@@

\paragraph{MT Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language
%\parcite{lopez2008statistical}.
Human evaluation is generally done by ranking the outputs of multiple systems
e.g., in the WMT tasks \parcite{bojar2015findings}, or by assigning
adequacy/fluency scores to each translation, a procedure recently improved
by \perscite{graham2015accurate}.
Some manual measures ask annotators to explicitly mark errors, which however
suffers from even lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.

@@

In order to address this shortcoming, more recent work quantified
the similarity of the reference and translation in terms
of their structure. \perscite{liu2005syntactic} took a syntactic approach, 
using dependency grammar, and
\perscite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
\perscite{gimenez2007linguistic} proposed to combine multiple types of information,
capturing the overlap between the translation and reference in terms of their
semantic (predicate-argument structures), lexical and morphosyntactic features.
\perscite{machacek:bojar:segranks:2015} divided the source sentences into shorter segments,
defined using a phrase structure parse, and applied human ranking to the resulting segments.\oa{is this description OK?}

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=0.35\textwidth]{ucca-tree-v2.png}

{\small
  \begin{tabular}{|l|l|l|l|}
\hline
L & Linker &  A & Participant \\
H & Parallel Scene & R & Relater \\
P & Process & C & Centre \\
\hline
\end{tabular}
}
    \end{center}
\caption{\label{fig:ucca_example_v2}
  Sample UCCA annotation. Leaves correspond to words and nodes to units.
  The dashed edge indicates that ``Tom'' is also a participant in the ``moved to NYC''
  Scene. Edge labels mark UCCA categories.}
%which are largely ignored in this work
\end{figure}

@@

UCCA's basic inventory of distinctions (its {\it foundational layer})
focuses on argument structures (adjectival, nominal, verbal and others)
and relations between them. The most basic notion is the Scene, which describes a movement, an
action or a state which persists in time. Each Scene contains one main
relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduation'' and ``moved''.
The participant ``Tom'' is a part of both Scenes, while ``NYC'' only of the
latter (\figref{fig:ucca_example_v2}). Further categories account for
inter-scene relations and the sub-structures of participants and relations.

@@

%Verbal argument structures are less stable in this sense, as they
%may be translated or paraphrased to a nominal or adjectival
%argument structure, and is thus less suitable as a basis for a semantic measure.
%E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
%but is similar in meaning to the one-verbed
%``after graduation, Tom moved to NYC''.

@@

Annotation of the semantic units requires first deciding whether
a unit is {\it structural}, i.e., has meaning-bearing sub-units also in the
target language, or {\it atomic}.
In most cases, atomic units
correspond to individual words, but they may also correspond to
multi-word expressions that translate as one unit. \oamod{For instance,
the expression ``took a shower'' is translated into the German ``duschte'',
while its individual words do not correspond to any sub-part of the German translation,
motivating the labeling the entire expression as an atomic node.}
When a multi-word unit is labeled as atomic, its sub-units' annotations are ignored
in the evaluation.

@@

%For example, consider the sentence ``After graduation, Tom moved
%to NYC'' as the source $s$, see \figref{fig:ucca_example}.
%The atomic units are the individual words,
%nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation $t$
%is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
%  be in English for presentation purposes. In our experiments, $t$ is in German,
%  Czech, Polish or Romanian.}.
%Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
%Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
%is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
%as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
%expresses the relation between the graduation and moving Scenes, is marked Bad,
%as their relation in $t$ is reversed.

\begin{figure}
    \begin{center}
    \includegraphics[width=0.35\textwidth]{ucca-tree-mteval-v2.png}
    \end{center}
  \caption{\label{fig:hume_tree_v2}
     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
     Atomic units are labelled using traffic lights (Red, Orange, Green) and structural
     units are marked A or B.}
\end{figure}
% Ondrej removes the following, he thinks that errors are not supposed to be
% propagated in HMEANT either:
%In HMEANT, it is not clear whether errors should be propegated up the SRL hierarchy.

Figure~\ref{fig:hume_tree_v2} presents an example of a HUME
annotation, where the translation is in English for ease of comprehension.
When evaluating ``to NYC'' the annotator looks at the translation and sees the
word ``stateside''. This word captures the whole phrase and so we mark this
non-leaf node with an atomic label. Here we choose Orange since
it approximately captures the meaning in this context.
The ability to mark non-leaves with atomic labels allows
the annotator to account for translations which only correspond at the phrase
level. Another feature highlighted in this example is that by separating structural
and atomic units, we are able to define where an error occurs, and localise
the error to its point of origin. The linker ``After'' is translated incorrectly as ``by''
which changes the meaning of the entire sentence. This error is captured at
the atomic level, and it is labelled Red. The sentence still contains two scenes and
a linker and therefore we mark the root node as structurally correct, Adequate.

@@

\figref{fig:interface} shows the HUME annotation interface\footnote{A demo of HUME
  can be found in \url{www.cs.huji.ac.il/~oabend/hume_demo.html}}.
% We now describe the HUME annotation interface, shown in \figref{fig:interface}. 
% %We turn to discuss the interface for collecting human judgements on the semantic units
% %of the source, relating to the translation. 
% UCCA annotations of the source are done offline,
% using the UCCA web application.
% %\figref{fig:interface} presents the user interface of HUME. 
% The translation appears at the top, and
% the source sentence directly below it. Underneath the source sentence appears
% its list of semantic units. 
% %For each semantic unit, including the one for the
% %entire source sentence, there is a textbox, displaying the yield of that unit.
% %Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
\oamod{One source sentence and one translation are presented at a time.}
The user is asked to select a label for each source semantic unit,
by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the right of the unit's box.
Units with multiple parents (as with ``Tom'' in \figref{fig:hume_tree_v2}) are displayed
twice, once under each of their parents, but are only annotatable in one of
their instances, to avoid double counting.

@@

The source sentences are all in English, and their UCCA annotation was performed by four
computational linguists and one linguist.
For the annotation of the MT output, we recruited two annotators for each of German, Romanian
and Polish and one main annotator for Czech. For Czech IAA,
several further annotators worked on a small number of 
sentences each. We treat these further annotators as one annotator, resulting in two annotators
for each language pair.
The annotators were all native speakers of the respective target languages and fluent in English.
\oa{maybe we should put in what training they've received}

@@

In order to assess the consistency of the annotation, we measure the Inter-Annotator
Agreement (IAA) using Cohen's Kappa on the multiply-annotated units.
%For German, Polish and Romanian we used two annotators, with a significant overlap
%between the sentences they annotated. For Czech we had the entire batch
%annotated by one main annotator and multiple secondary annotators, 
%each annotating a sub-part of the batch.
\tabref{tab:iaa} reports the number of units which have two annotations from
different annotators and the corresponding Kappas.
%
%In order to compute Kappa, we consider the annotation task as a classification task on 
%nodes, and compute agreement on nodes which have two annotations from different annotators.
We report the overall Kappa, as well as separate Kappas on atomic
units (annotated as Red, Orange or Green) and structural units (annotated
as Adequate or Bad).
As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
is generally little confusion between the two types of units.

@@

\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_heatmap_#1.png}}

@@

@@

\oa{This entire section needs reviewing}

We discuss the main differences between HUME and HMEANT,
a human MT evaluation metric that measures the overlap between the translation a reference
in terms of their SRL annotations. The discussion is acoompanied by examples from our
experimental data.

@@

Conversely, HMEANT does not penalize errors (e.g., incorrect tense or negation flip) in a correctly aligned predicate:

\begin{center}
\begin{tabular}{lp{5.4cm}}
Source & \small ... don't skip meals ... \\
Transl.& \small ... nicht {\"u}berspringen Mahlzeiten ... \\
Gloss & \small ... don't skip meals ... \\
\end{tabular}
\end{center}

While the translation is mostly understandable, the verb {\"u}berspringen is incorrectly inflected. HUME allows marking the main verb as Orange, while
marking the translation of the argument (``Malhzeiten'' or ``meals'') as correct. HMEANT may only mark the predicate as correct, or mark the entire argument structure as an error.

It is also relatively common that verbal constructions are trasnlated
as non-verbal ones or vice versa. Consider the following example:

>>>>

HUME operates by aggregating human assessments of the translation quality of individual
semantic units in the source sentence. We 
are thus avoiding the semantic annotation of machine-generated text,
which is often garbled or semantically unclear.
This also allows the re-use of the source semantic annotation for
measuring the quality of different translations of the same source sentence
and avoids relying on reference translations, 
which have been shown to bias annotators~\cite{fomicheva-specia_ACL:2016}.

@@

\paragraph{MT Evaluation.}
%MT evaluation has been traditionally pursued through human-based
%evaluation that quantifies the adequacy of the translation in preserving the meaning 
%of the source sentence, and its fluency as a sentence in the target language
%\parcite{lopez2008statistical}.
Human evaluation is generally done by ranking the outputs of multiple systems
e.g., in the WMT tasks \parcite{bojar2015findings}, or by assigning
adequacy/fluency scores to each translation, a procedure recently improved
by \perscite{graham2015accurate} and called direct assessment. We use this method to compare and contrast
with HUME later in the paper. HTER~\parcite{snover2006study} is another widely used human evaluation metric
which uses edit distance metrics to compare a translation and its human 
post-edition. HTER suffers from the problem that small edits in the translation
could in fact be serious flaws in accuracy, eg. deleting a negation. 
Some manual measures ask annotators to explicitly mark errors, which however
suffers from even lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.

@@

In order to address this shortcoming, more recent work quantified
the similarity of the reference and translation in terms
of their structure. \perscite{liu2005syntactic} took a syntactic approach, 
using dependency grammar, and
\perscite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
\perscite{gimenez2007linguistic} proposed to combine multiple types of information,
capturing the overlap between the translation and reference in terms of their
semantic (predicate-argument structures), lexical and morphosyntactic features.
\perscite{machacek:bojar:segranks:2015} divided the source sentences into shorter segments,
defined using a phrase structure parse, and applied human ranking to the resulting segments.

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=0.35\textwidth]{ucca-tree-v3.png}

{\small
  \begin{tabular}{|l|l|l|l|}
\hline
L & Linker &  A & Participant \\
H & Parallel Scene & R & Relater \\
P & Process & C & Centre \\
\hline
\end{tabular}
}
    \end{center}
\caption{\label{fig:ucca_example_v2}
  Sample UCCA annotation. Leaves correspond to words and nodes to units.
  The dashed edge indicates that ``Tom'' is also a participant in the ``moved to America''
  Scene. Edge labels mark UCCA categories.}
%which are largely ignored in this work
\end{figure}

@@

UCCA's basic inventory of distinctions (its {\it foundational layer})
focuses on argument structures (adjectival, nominal, verbal and others)
and relations between them. The most basic notion is the Scene, which describes a movement, an
action or a state which persists in time. Each Scene contains one main
relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to America''
contains two Scenes, whose main relations are ``graduation'' and ``moved''.
The participant ``Tom'' is a part of both Scenes, while ``America'' only of the
latter (\figref{fig:ucca_example_v2}). Further categories account for
inter-scene relations and the sub-structures of participants and relations.

@@

%Verbal argument structures are less stable in this sense, as they
%may be translated or paraphrased to a nominal or adjectival
%argument structure, and is thus less suitable as a basis for a semantic measure.
%E.g., ``after he graduated, Tom moved to America'' contains two verbs,
%but is similar in meaning to the one-verbed
%``after graduation, Tom moved to America''.

@@

Annotation of the semantic units requires first deciding whether
a unit is {\it structural}, i.e., has meaning-bearing sub-units also in the
target language, or {\it atomic}.
In most cases, atomic units
correspond to individual words, but they may also correspond to
multi-word expressions that translate as one unit. For instance,
the expression ``took a shower'' is translated into the German ``duschte'',
while its individual words do not correspond to any sub-part of the German translation,
motivating the labeling the entire expression as an atomic node.
When a multi-word unit is labeled as atomic, its sub-units' annotations are ignored
in the evaluation.

@@

%For example, consider the sentence ``After graduation, Tom moved
%to America'' as the source $s$, see \figref{fig:ucca_example}.
%The atomic units are the individual words,
%nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation $t$
%is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
%  be in English for presentation purposes. In our experiments, $t$ is in German,
%  Czech, Polish or Romanian.}.
%Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
%Orange (as it corresponds to ``will move'' in $t$), and ``America''
%is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
%as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
%expresses the relation between the graduation and moving Scenes, is marked Bad,
%as their relation in $t$ is reversed.

\begin{figure}
    \begin{center}
    \includegraphics[width=0.35\textwidth]{ucca-tree-mteval-v3.png}
    \end{center}
  \caption{\label{fig:hume_tree_v2}
     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
     Atomic units are labelled using traffic lights (Red, Orange, Green) and structural
     units are marked A or B.}
\end{figure}
% Ondrej removes the following, he thinks that errors are not supposed to be
% propagated in HMEANT either:
%In HMEANT, it is not clear whether errors should be propegated up the SRL hierarchy.

Figure~\ref{fig:hume_tree_v2} presents an example of a HUME
annotation, where the translation is in English for ease of comprehension.
When evaluating ``to America'' the annotator looks at the translation and sees the
word ``stateside''. This word captures the whole phrase and so we mark this
non-leaf node with an atomic label. Here we choose Orange since
it approximately captures the meaning in this context.
The ability to mark non-leaves with atomic labels allows
the annotator to account for translations which only correspond at the phrase
level. Another feature highlighted in this example is that by separating structural
and atomic units, we are able to define where an error occurs, and localise
the error to its point of origin. The linker ``After'' is translated incorrectly as ``by''
which changes the meaning of the entire sentence. This error is captured at
the atomic level, and it is labelled Red. The sentence still contains two scenes and
a linker and therefore we mark the root node as structurally correct, Adequate.

@@

\figref{fig:interface} shows the HUME annotation interface\footnote{A demo of HUME
  can be found in \url{www.cs.huji.ac.il/~oabend/hume_demo.html}}.
% We now describe the HUME annotation interface, shown in \figref{fig:interface}. 
% %We turn to discuss the interface for collecting human judgements on the semantic units
% %of the source, relating to the translation. 
% UCCA annotations of the source are done offline,
% using the UCCA web application.
% %\figref{fig:interface} presents the user interface of HUME. 
% The translation appears at the top, and
% the source sentence directly below it. Underneath the source sentence appears
% its list of semantic units. 
% %For each semantic unit, including the one for the
% %entire source sentence, there is a textbox, displaying the yield of that unit.
% %Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
One source sentence and one translation are presented at a time.
The user is asked to select a label for each source semantic unit,
by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the right of the unit's box.
Units with multiple parents (as with ``Tom'' in \figref{fig:hume_tree_v2}) are displayed
twice, once under each of their parents, but are only annotatable in one of
their instances, to avoid double counting.

@@

The source sentences are all in English, and their UCCA annotation was performed by four
computational linguists and one linguist.
For the annotation of the MT output, we recruited two annotators for each of German, Romanian
and Polish and one main annotator for Czech. For Czech IAA,
several further annotators worked on a small number of 
sentences each. We treat these further annotators as one annotator, resulting in two annotators
for each language pair.
The annotators were all native speakers of the respective target languages and fluent in English.
They completed a three hour on-line training session where we described UCCA and the HUME task and then
we worked through a number of examples.

@@

In order to assess the consistency of the annotation, we measure the Inter-Annotator
Agreement (IAA) using Cohen's Kappa on the multiply-annotated units.
%For German, Polish and Romanian we used two annotators, with a significant overlap
%between the sentences they annotated. For Czech we had the entire batch
%annotated by one main annotator and multiple secondary annotators, 
%each annotating a sub-part of the batch.
\tabref{tab:iaa} reports the number of units which have two annotations from
different annotators and the corresponding Kappas.
%
%In order to compute Kappa, we consider the annotation task as a classification task on 
%nodes, and compute agreement on nodes which have two annotations from different annotators.
We report the overall Kappa, as well as separate Kappas on atomic
units (annotated as Red, Orange or Green) and structural units (annotated
as Adequate or Bad).
As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
is generally little confusion between the two types of units.
This results in the Kappa for all units being considerably higher than the kappa
over the atomic units or structural units, where there is more internal confusion.

@@

\def\iaafig #1{\includegraphics[width=4cm]{iaa_heatmap_#1.png}}

@@

We have just presented a quantitative comparison between DA and HUME. DA is however, a very
different type of metric to HUME. It returns just one score per sentence and requires annotators to make 
decisions collating many smaller judgements about individual errors in the sentence. 
Comparing DA and HUME is important in order to see where they agree and differ in order to
determine under what circumstances it would be appropriate to apply each method. 
In cases where we specifically want to know about the quality of semantic
components in a translation, HUME and HMEANT could be considered.

@@

HMEANT is a human MT evaluation metric that measures the overlap between the translation a reference
in terms of their SRL annotations.
In this section we present a qualitative comparison between HUME and HMEANT
and the discussion is accompanied by examples from our
experimental data.

@@

%This is not entirely true - they do allow an alignment with partial
%Conversely, HMEANT does not penalize errors (e.g., incorrect tense or negation flip) in a correctly aligned predicate:
%
%\begin{center}
%\begin{tabular}{lp{5.4cm}}
%Source & \small ... don't skip meals ... \\
%Transl.& \small ... nicht {\"u}berspringen Mahlzeiten ... \\
%Gloss & \small ... don't skip meals ... \\
%\end{tabular}
%\end{center}
%While the translation is mostly understandable, the verb {\"u}berspringen is incorrectly inflected. HUME allows marking the %main verb as Orange, while
%marking the translation of the argument (``Malhzeiten'' or ``meals'') as correct. HMEANT may only mark the predicate as %correct, or mark the entire argument structure as an error.

It is also relatively common that verbal constructions are translated
as non-verbal ones or vice versa. Consider the following example:
